---
title: "mice"
output: html_document
date: "2025-03-11"
---
### Mice imputation for ideology

We look at the NA distribution to see which variables have more missing values.

```{r}
colSums(is.na(final_data))
```

It can be seen that the ideology variable is the most appropriate variable to impute as its number of NA represents a significant percentage of the total observations.

```{r}
mice_data <- final_data |> 
  mutate(across(where(is.labelled), as_factor),
         ideology = as.numeric(ideology)) |>
  select(-caseid, -religiosity_percent) # not useful for the mice imputation
```

There are factor and numerical variables, so the best option is to test with Random Forest and Pmm methods.

```{r}
imputed_comparison <- data.frame(
  original = mice_data$ideology,
  imputed_rf = complete(mice(mice_data, method = "rf", 
                             m = 3, maxit = 3, 
                             seed = 123))$ideology,
  imputed_pmm = complete(mice(mice_data, method = "pmm",
                              m = 3, maxit = 3, 
                              seed = 123))$ideology)
```

The m and the maxit arguments are = 3 to reduce loading time without losing so much information.

The next step is to plot the results to compare the resulting distributions: 

```{r}
# Arguments for the plot 
methods <- c("original", "imputed_rf" , "imputed_pmm")
titles <- c("Distribution of the Age Variable",
           "Random Forest-imputed Distribution",
            "PMM-imputed Distribution")
colors_fill <- c( "#E7CB94", "#E7969C", "#DE9ED6")


# Long format 
data_imputed_long <- imputed_comparison |>
  pivot_longer(cols = all_of(methods), names_to = "method", 
                values_to = "value") |>
  mutate(title = factor(method, levels = methods, labels = titles))


# Distributions comparison 
plot_mice <- ggplot(data_imputed_long, aes(x = value, fill = title)) +
  geom_histogram(binwidth = 1, color = "black", position = "identity") +
  facet_wrap(~ title, scales = "free_y") +
  scale_fill_manual(values = colors_fill) +
  theme_classic() +
  theme(legend.position = "none")

plot_mice
```


```{r}
init = mice(mice_data)
```

```{r}
imputed_data <- complete(init)

write.csv(imputed_data, "imputed_data.csv")
# my_data <- read.csv("my_data.csv")
```

```{r}
meth = init$method

meth

meth[c("trans_name")]=""

imputed_data_final = mice(mice_data, method=meth, m=5, seed=123)
```
```{r}
summary(imputed_data_final)
```


```{r}
imputed_data_final <- complete(imputed_data_final)

write.csv(imputed_data_final, "imputed_data_final.csv")
# my_data <- read.csv("my_data.csv")
```

# models 
```{r}
imputed_data_final_2 <- imputed_data_final_2 |> 
  mutate(gdp_pc = log(gdp_pc))

imputed_data_final_2 <- imputed_data_final_2 |> 
  mutate(isocntry = factor(isocntry))

imputed_data_final_2$contact_lgbti <- relevel(imputed_data_final_2$contact_lgbti, ref = "No contact")


imputed_data_final_2$age <- as.numeric(as.character(imputed_data_final_2$age))
# mputed_data_final_2$ideology <- as.numeric(as.character(imputed_data_final_2$ideology))
imputed_data_final_2$self_determination <- as_factor(imputed_data_final_2$self_determination)

imputed_data_final_2$personal_satis <- as.factor(imputed_data_final_2$personal_satis)
imputed_data_final_2$social_class <- as.factor(imputed_data_final_2$social_class)


```


```{r}
model <- glm(trans_name ~ age + 
               female + 
               religion + 
               occupation + 
               marital_status + 
               personal_satis +
               contact_lgbti +
               ideology +
               social_class +
               gdp_pc +
               self_determination + 
               gini + 
               rain_ind, 
             family = "binomial", 
             data = imputed_data_final_2)

summary(model)


model2 <- glmer(trans_name ~ 
                  age + 
                  female + 
                  religion +
                  occupation +
                  marital_status + 
                  personal_satis +
                  contact_lgbti +
                  ideology +
                  social_class +
                  (1|isocntry), 
                family = "binomial", 
                data = imputed_data_final_2)

summary(model2)


```

```{r}
final_model <- glmer(trans_name ~ 
                       scale(age) + # Cuando incluyes tanto age como age^2, 
                       #el término lineal (age) y el término cuadrático (age^2) trabajan
                       # juntos para modelar una relación curvilínea, entonces aunque age no es significativa 
                       # no la vamos a eliminar
                       I(scale(age)^2) +
                       female + 
                       occupation +
                       religion +
                       personal_satis +
                       contact_lgbti*scale(rain_ind) +
                       self_determination +
                       scale(ideology)*scale(gdp_pc) +
                       (1  + scale(age) +
                          female +
                          scale(ideology)
                        |isocntry), 
                     family = binomial(link = "logit"),
                     control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
                     data = imputed_data_final_2)

str(imputed_data_final_2)

summary(final_model)
```

```{r}
# Logistic prediction (no va)

set.seed(123)

wofi <- imputed_data_final_2 |> 
  filter(isocntry != "FI")

test_country <-imputed_data_final_2 |> 
  filter(isocntry == "FI")

train_set <- wofi %>%
  group_by(isocntry) %>%
  sample_frac(0.7) %>%
  ungroup()

test_set <- anti_join(wofi, train_set, by = c("serialid")) 

testing <- bind_rows(test_set, test_country)

train_set <- train_set |> 
  select(trans_name:religiosity_percent, isocntry)

test_set <- test_set |> 
  select(trans_name:religiosity_percent, isocntry)

library(caret)
library(DataExplorer)
plot_intro(imputed_data_final_2) 

imputed_data_final_3 <- imputed_data_final_2 %>% 
  filter(!is.na(trans_name))

write.csv(imputed_data_final_3, "imputed_data_final_3.csv")
# my_data <- read.csv("my_data.csv")

in_train <- createDataPartition(imputed_data_final_3$trans_name, p = 0.7, list = FALSE)  # 70% for training
training <- imputed_data_final_3[ in_train,]
testing <- imputed_data_final_3[-in_train,]

final_model <- glmer(trans_name ~ 
                       scale(age) + # Cuando incluyes tanto age como age^2, 
                       #el término lineal (age) y el término cuadrático (age^2) trabajan
                       # juntos para modelar una relación curvilínea, entonces aunque age no es significativa 
                       # no la vamos a eliminar
                       I(scale(age)^2) +
                       female + 
                       occupation + 
                       religion +  # Si la interacciín es significativa no debemos quitar las variables individuales aunque no lo sean
                       personal_satis +
                       contact_lgbti*scale(rain_ind) +
                       self_determination +
                       scale(ideology)*scale(gdp_pc) +
                       (1  + scale(age) +
                          female +
                          scale(ideology)
                        |isocntry), 
                     family = binomial(link = "logit"),
                     control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
                     data = training)

summary(final_model)

train <- training[complete.cases(training), ]

test <- testing[complete.cases(testing), ]


colSums(is.na(testing))  # Check for missing values in testing


pred <- predict(final_model, newdata = test, type = "response")

prediction <- as.factor(ifelse(pred>0.5, 1, 0))
prediction <- factor(ifelse(pred > 0.5, 1, 0), levels = c(0, 1))

length(prediction)

str(test)
str(prediction)

attr(prediction, "names") <- NULL


levels(prediction)
levels(test$trans_name)


test$trans_name <- factor(test$trans_name, levels = c(0, 1))

confusionMatrix(prediction, test$trans_name)$table
confusionMatrix(prediction, test$trans_name)$overall[1:2]

str(x)

```

